\chapter{Implementation details}
\label{chapter:implementation-details}

\section{Comunication between QGroundControl and FlightGear Flight Simulator}

Because QGroundControl uses the special \textit{heartbeat} message another
communication, different from the already provided module. FlightGear is able 
to send messages that contain only telemetry data and other UAV specific parameters, 
thus making the creating of a \textit{heartbeat} message impossible.

One approach for overcoming this shortcoming would be the development and 
integration of a \textit{proxy} program that forwards to QGC the messages
received from the FGFS instances. The proxy would also emit heartbeat messages
at regular intervals of time. The drawback of this approach is represented
by the fact that a communication delay would be introduced by this extra module.
Another approach would be the implementation of a custom communication protocol, 
embedded in QGC that is not based on a specific \textit{heartbeat} message.

For this thesis the MAVLink protocol was bypassed and we implemented a generic
protocol based on the messages sent by FGFS. Qt's signal mechanism permitted
event generation every time a message was received from FGFS.
Two UDP sockets (one for input and one for output) for each UAV are used to
ensure communication between QGC and FGFS. 
Every time a message is received from FGFS it is parsed according to the XML 
configuration used at the startup of FGFS and the position, speed, orientation
and other parameters are set and the MAV's position is rendered on the map.
In the proposed approach the data received from FGFS is manually parsed and 
any message is considered to be a heartbeat message.

\section{Formation Flight Module Implementation}
\label{sec:code}

The formation flight module was developed using the architecture describe in 
\labelindexref{Section}{sec:architecture}.

The connection between FGFS instances and QGC, the later was modified to establish
UDP connection. For each monitored FGFS instance,  QGC opens another UDP socket
on which it will send data back. For communication with FGFS, QGC uses a total
number of $n+1$ sockets, n representing the number of FGFS instances and the last
one is the UPD input socket. Also, I have modified the communication used
by a MAV so that it does not use MAVLink. Instead it uses a generic protocol
based on the FGFS protocol XML. Every time a message is received it is parsed
and the position, orientation, speed and other telemetry data are set in 
QGC; After this data has been set a refresh signal is emitted and the position
on the map is updated for each UAV.

Another QGC widgets was also developed in the \textit{Autonomous UAV} project
where the mission details can be configured. Using this widget a human user
can select the number of UAVs used, their flight path, the equipment for every
drone and the leader of the formation.

The formation flight module is written in C++ using the Boost library. The 
module connects to the CAN interface simulator via a TCP socket. When the module
will be introduced on the Hirrus drone, on the Raspberry PI a character device
will be crated that will read data from a CAN bus. Each drone will be equipped
with a Raspberry PI that will run a Mission Supervisor.

At startup the module reads a JSON configuration file. The configuration
file contains the number of drone and the geometry formation. Each UAV has
a relative altitude to another UAV, a position (left or right) to the possible 
leader and at what distance it should be from the other UAVs. For extensibility
reasons the JSON also contains data about the equipment. An example of a formation
configuration file can be seen in \labelindexref{Listing}{lst:formation-json}

After reading the JSON file, the data is store in a data structure called 
\textbf{uav_formation}. The \textit{uav_formation} class contains the following fields
and methods:

\begin{description}
\item [has_leader] \hfill \\ boolean flag that specifies if the formation has a UAV leader
or a virtual leader
\item [name] \hfill \\ the name of the formation; eg: V Formation
\item [drones] \hfill \\ a vector of uav_drone structures that contain data about all the
UAVs in formation.
\item [getLeader] \hfill \\ returns the uav_drone if the formation has a UAV leader
\item [getDrone] \hfill \\ returns the uav_drone with the desired name
\end{description}

An \textbf{uav_drone} class contains the configuration for each drone of how it
should behave in formation and has the following fields:

\begin{description}
\item [name] \hfill \\ the name of the UAV drone; eg: MAV 01
\item [role] \hfill \\ specifies the role based on the UAV equipment; eg: Infrared Scanner
\item [personal_space] \hfill \\ radius for a sphere were if another UAV is detected, mild
evasive actions should be taken to avoid collision;
\item [emergency_space] \hfill \\ radius for a shpere where if another UAV is detected, 
the collision avoidance module will have maximum priority
\item [is_leader] \hfill \\ boolean marker that specifies if the current drone acts
as the leader of the formation
\item [geometry] \hfill \\ array of uav_geometry specifying the relative position of the 
current UAV to other UAVs
\end{description}

The \textbf{uav_geometry} contains data about the relative position to another drone and
it contains the following fields:

\begin{description}
\item [relative_drone] \hfill \\ the UAV to which the geometry is relative to
\item [distance] \hfill \\ array of three values containing the height difference
between the UAVs, the heading multiply factor (used for determining how fast a UAV
will turn to enter formation), the distance between the UAVs
\end{description}

The module starts to send commands to the CAN Interface as soon as initial 
position data is available about all the UAVs. Based on the number of drones
needed in formation, the module waits until all the necessary data is received.
As long as the initial data is not present, the UAV flight is not influenced 
by this module.

Data about each UAV is kept in a data structure called \textbf{position_container}.
The \textit{position_container} acts as a hash map, keeping as key the name 
of the UAV drone and as value a pair of two \textit{uav_position} structures.
The two \textit{uav_position} instances represent the current and previous
positions for each aircraft.

The \textbf{uav_position} class is the main class of the project and is
responsible with creating commands for Flight Gear Instance. The class
contains the following fields:

\begin{description}
\item [id] \hfill \\ The id/name of the UAV
\item [latitude] \hfill \\ The GPS latitude
\item [longitude] \hfill \\ The GPS longitude
\item [altitude] \hfill \\ The altitude difference from sea level
\item [heading] \hfill \\ The true heading that the UAV drone is following
\item [speed] \hfill \\ The instantaneous speed of the UAV at the time of reporting
\item [x] \hfill \\ Value on ECEF x-axis
\item [y] \hfill \\ Value on ECEF y-axis
\item [z] \hfill \\ Value on ECEF z-axis
\end{description}
% 
This class is also responsible for converting the GPS coordinates to ECEF coordinates
and back. These methods are called: \textit{lla2ecef} and \textit{ecef2lla}. In 
this class the static method \textit{parse_uav} creates a uav_position instance
from a message received from the CAN Simulator. The received message
is parsed using a regular expression and the position, GPS position,  altitude, 
altitude heading and speed. After the ECEF data is set using the method 
\textit{lla2ecef}.
 
Based on the formation configuration and current position, a command is created 
by setting the same altitude as the leader and determining the heading 
from the current position and and the desired position. For example, if the UAV
is on the right of the leader and has to be on his left, the UAV receives a
true heading that is smaller than the leaders heading. The UAV maintains this heading
until the distance from the leader is in a desired range and is in the correct
position. On the other hand if the UAV is on the left side of the leader and
the formation specifies that it has to be on the left side, but the distance
from the leader is to big, a true heading that is greater than the leader is 
reported, until the distance is in the desired range. To determine the side on
which the UAV resides,  from the two known position of the leader 
the flight direction is calculated and if the point is on the left or right.
In this calculations the altitude is ignored and considering that $A$ is the 
leaders previous location and $B$ is the leaders current location and $C$
is the point for which the side is computed, the formula used is the 
cross-product is \labelindexref{Equation}{eqn:side-of-line}:

\begin{equation}
sign = \begin{vmatrix}B.x - A.x & B.y - A.y \\ C.x - A.x & C.y - A.y\end{vmatrix}
\qquad\parbox{4.0cm}{\footnotesize$\begin{aligned} 
  sign > 0  & \text{ point on left side of line}
  \\[-1.0ex] sign = 0  & \text{ point on same line}
  \\[-1.0ex] sign < 0  & \text{ point on right side of line}
  \end{aligned}$}
\label{eqn:side-of-line}
\end{equation}

In the conversion between GPS and ECEF, a computational error is introduced 
in the mathematical model,  mostly because the formulas consider the Earth Radius
of constant size. In a \textit{Line Astern} formation, this error determines
the agents to slightly move from left to right. Also this error can be seen
in calculating the distance between two UAVs. The distance computed in 
ECEF coordinates system using the Euclidean formal varies from the distance 
computed in GPS space using the Haversine formula with values ranging between
10 and 30 meters,  for distances between 100 and 200 meters.